---
title: "Vertical distribution of the phytoplankton in the water column in the Baffin Bay"
description: |
  Presentation of the main results.
author:
  - name: Philippe Massicotte
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    theme: theme.css
bibliography: /home/filoche/Documents/library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "center"
)

library(here)
library(pdftools)
library(data.table)
library(tidyverse)
```

# Context

This project aims at exploring how various metrics/proxies can give insights on the vertical distribution of the phytoplankton biomass in the water column using the data measured during the Green Edge oceanographic cruise conducted in the Baffin Bay in 2016. *Four main indicators were considered to get insights about the phytoplankton biomass in the water column:*

1.  Fluorescence
2.  Pigments
3.  Particle beam attenuation coefficient (CP)
4.  Particle backscattering coefficient (bbp)

Most of the data is presented as a function of open water days (OWD) that were calculated in @Randelhoff2019. The first graph shows the OWD of each station.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_map_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Data processing for 3D plots

During the meeting we had on 2021-01-22, it was questioned why I was averaging data by `owd` and `depth` before doing the interpolation. This is simply because it is possible that there are many observations for a specific pair of `owd` and `depth`. For example, it can be seen here that there are 9 values (i.e. 9 stations) that have an `owd = 2` and `depth = 1.980`.

```{r count-duplicated-owd-depth, cache=TRUE}
ctd <- fread(here::here("data/clean/ctd.csv"))

ctd[, .N, by = .(owd, depth_m)][N > 5] %>% 
  as_tibble() %>% 
  arrange(desc(N)) %>% 
  head(6) %>% 
  knitr::kable()
```

Here are 9 stations in the CTD data with `owd = 2` and `depth = 1.980`.

```{r count-duplicated-owd-depth2, cache=TRUE}
ctd %>%
  as_tibble() %>%
  filter(owd == 2 & depth_m == 1.98) %>%
  select(station, depth_m, owd, flor_mg_m3) %>%
  rmarkdown::paged_table()
```

Hence, before performing the interpolation, data has to be averaged by OWD and depth.

# Fluorescence

This section shows the vertical distribution of phytoplankton based on fluorescence.

## Fluorescence from the CTD

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_phytoplankton_biomass_ctd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

Using boxplots can be one interesting alternative way to present the data. For example, I have divided the CTD fluorescence data presented in the above graphic into ice-free/ice-covered and above/below the 0.1 isolume.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_boxplot_ctd_fluorescence_isolume_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## Fluorescence from the MVP

*Note that because the MVP is measuring continually, there are no stations associated with each measurement. This is why that the MVP data is not presented as a function of OWD.*

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/05_phytoplankton_biomass_mvp_fluo.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Pigments

These graphs were made using the pigments data from the rosette. Pigments were summed into two groups:

1. Photoprotection pigments
2. Accessory pigments

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/04_phytoplankton_biomass_pigments.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Particle beam attenuation coefficient (CP)

> The increase of cp during the lighted portion of the day is usually explained by the accumulation of intracellular carbon concentration associated with photosynthetic processes [@Kheireddine2014].

## CP from the CTD

The CP visualization based on the CTD can be seen above. There is a clear relationship between fluorescence and CP as seen below.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_scatterplot_fluorescence_vs_cp.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

The next graph shows the same data but divided differently. The best relation can be seen for the observations below the 0.1 isolume and located in open water. *I think the bottom left panel is the most interesting where **we can see the L shape**.*

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_scatterplot_fluorescence_vs_cp_isolume_depth.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

Finally, the same data by transect.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/02_scatterplot_fluorescence_vs_cp_by_transect.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## CP from the MVP

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/05_phytoplankton_biomass_mvp_cp.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Particle backscattering coefficient (bbp)

## bbp from the hydroscat

These graphs show bbp at six different wavelengths measured by the hydroscat.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/03_phytoplankton_biomass_hydroscat.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Primary production

## in-situ primary production

Visualization of in-situ primary production based on the data from Patrick Rimbault.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/07_insitu_primary_production.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## Estimated primary production

A lot of work has been done for this section. *The goal was to calculate primary production based on photosynthetic parameters.* Note that estimation of daily primary production was calculated between the surface and the isolume depth corresponding to the 0.1 mol photons $m^{-2} d^{-1}$.

This is the list of data used to estimate primary production:

1. Photosynthetic parameters
2. Incident PAR calculated from SBDART.
3. PAR transmittance in the water column calculated from the C-OPS.
4. Chla from the CTD.

The next graph shows these intermediate parameters used to estimate primary production for four different stations.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/16_02_stations_with_highest_daily_pp.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

Finally, we can visualize estimated primary production as a function of OWD.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/20_02_primary_production_pvse_vs_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## Comparing in-situ and estimated primary production

Interestingly, we can see that there is a good relationship between measured and estimated primary production (considering that both in-situ and estimated values are independent). We can also observe that at low values, estimated primary production is somehow underestimated compared to in-situ measurements.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/21_01_scatterplot_primary_production_insitu_vs_pvse.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

To get more insights, I have divided the data into three main categories: Amundsen, barge and ice-covered stations.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/21_01_scatterplot_primary_production_insitu_vs_pvse_by_deployment.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Particulate organic carbon (POC)

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/18_poc_vs_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

# Phytoplankton absorption

Maybe we could use absorption information to get some additional insights.

## Vertical profiles

This graph shows the vertical profiles of phytoplankton absorption at 440 nm. It seems that open water stations have subsurface maximum more apparent than the profiles of the ice-covered stations.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/22_vertical_profiles_aphy.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## Spectral profiles

By looking at the spectral profiles, we can also notice that there are indeed differences in the water column. 

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/22_spectral_profiles_phyto_absorption_isolume_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

Finally, I have averaged the spectral profiles in four categories. I think these can be seen as end-members spectra.

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/22_spectral_profiles_phyto_absorption_averaged_isolume_owd.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

## Comparing blue vs red regions

With all the data

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/23_scatterplot_phytoplankton_specific_absorption_blue_red.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```

By groups

```{r}
tf <- tempfile(fileext = ".png")
res <- pdf_convert(
  here("graphs/23_scatterplot_phytoplankton_specific_absorption_blue_red_by_group.pdf"),
  filenames = tf,
  format = "png",
  dpi = 200,
  verbose = FALSE,
  antialias = TRUE
)
knitr::include_graphics(tf)
```
